---
title: "Assignment 1: Navigating R Studio"
---

This will be much more usual friendly if you go through this file is 'Visual' mode! Press the 'Visual' button above (it should be RIGHT above the "title" of this document).

# Navigating RStudio

This is an **R Notebook**. When you execute code within the notebook, the results appear beneath the code. Try executing this chunk by clicking the *Run/Play* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```{r}
2 + 3
```

Now try executing a chunk with multiple equations. Feel free to edit inside it to play around! Simple click inside the code chunk and type like a normal word document. Play around with the different equations!

```{r}
2 + 3
2 - 3
2 * 3
2 / 3
2 ^ 3
```

You are able to write freely when outside of a code chunk. However, within a code chunk, R sees everything as code and won't know how to process random sentences and words. To comment inside a code chunk, simply include a *\#* beforehand. Comments are incredibly helpful for both you and the viewer of your code, and we encourage you to get in the habit of annotating your code as you go!

Try running this code chunk below. Then, delete the *\#* and notice the error that occurs.

```{r}
# This is a comment. Below is an equation to solve for 5 minus 2.
5 - 2
```

### Try creating a code chunk yourself!!

Add a new chunk below this by clicking the *Code -\> Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*. Inside your chunk, (1) create a comment describing your current feelings towards R. Next, think of a very simple equation, and (2) create a comment describing what the equation will solve and then (3) write this equation out and run it.

# Working with variables

Now that you have a better sense of navigating R and using basic arithmetics- lets jump into creating variables! Let's start with a basic variable named my_apples. Run this code, and notice in your environment (top right window of RStudio) that a variable my_apples has been added! *Note: that variables cannot have invalid characters, such as spaces. Try replacing the underscore with a space and seeing the error that follows!*

```{r}
# Assinging the value of 5 to my_apples
my_apples <- 5
```

Additionally, if you ever want to remind yourself what my_apples is, you can print my_apples using code and have the output tell you the value. Try running this code below.

```{r}
# Print the value of my_apples
my_apples
```

But uh oh! You got hungry and ate an apple. How can we represent that? Well, there are multiple different ways we could do that. The first is simply reassigning my_apples to the number 4. Run this code, and notice in the environment

```{r}
# Assign the value of 4 to my_apples
my_apples <- 4
# Print the value of my_apples
my_apples
```

While this works, it is not very practical. Imagine you were working with very large values and a large number of variables. Manually reassigning every variables would be a terrible idea. So instead, you can actually assign a variable to itself with basic math like we did above. Here is an example.

```{r}
# Reassigning the value of 5 to my_apples
my_apples <- 5

# Print the current value of my_apples
my_apples

# Assign my_apples to whatever my_apples was minus 1
my_apples <- my_apples - 1

# Print the new value of my_apples
my_apples
```

You can even do math using only variables instead of numbers. For example, let's double the amount of apples we have. (Just like in real math, you can do this by multiplying or adding the value by itself. For this example, we are adding.)

```{r}
# Reassigning the value of 5 to my_apples
my_apples <- 5

# Double the amount of apples
my_apples <- my_apples + my_apples

# Print the new value of my_apples
my_apples
```

Note: you won't always have to reassign the value of 5 to my_apples. We are only doing it here in case you run any chunks out of order. But at this point, we are going to show you a couple of examples. The first is the case when you simply tell R to print my_apples with math. Run the code a few times and notice the output.

```{r}
# Print my_apples
my_apples

# Print my_apples minus 1
my_apples - 1
```

Spoiler: it never changed! That is because you never changed the value of my_apples- you simply told R to output the number minus the value of 1! However, if you did reassign the values, note it will continuously decrease by the amount. Run this next code chunk a few times to see for yourself.

```{r}
# Print my_apples
my_apples

# Assign my_apples to my_apples minus 1
my_apples <- my_apples - 1

# Print the new value of my_apples
my_apples
```

## Practice with variables

Try it yourself! For this, we are going to explicitly tell you what to write, and you will be in charge of filling in the code chunks underneath our comments.

(1) **Let's create a new variable titled "my_oranges". We will assign the value 3 to oranges.**

```{r}
# Create a new varaible my_oranges with a value of 3 below

```

(2) **Notice that my_apples should still be in your environment after running the above code! Just in case you ran any of the chunks out of order, we are going to remind R what "my_apples" is. Your job now is to create a new variable titled "total_fruit"**

```{r}
# Reminding R what apples is. Run this BEFORE the next code chunk!
my_apples <- 5
```

```{r}
# Create a new variable called total_fruit taking apples plus oranges

```

(3) **Now let's say you ate 2 apples. Below, subtract my_apples by two, and then execute the total_fruit variable above. Notice how it uses the latest version of my_apples as opposed to the value listed above. In R, variables are always assigned based on the last code run. Feel free to rerun this code a few times, and reset with the my_apples \<- 5 above anytime you'd like.**

```{r}
# Subtract your number of apples by 2

```

# Data types

### Recap of data types in R

-   Decimals values like `4.5` are called **numerics**.
-   Natural numbers like `4` are called **integers**. Integers are also numerics.
-   Boolean values (`TRUE` or `FALSE`) are called **logical**.
-   Text (or string) values are called **characters** `"character"` .
-   **Complex** is also a data type- we will not be discussing that further at the moment.

Characters are surrounded by quotation marks. You can either use " or ' as quotation marks actually, as long as they are consistent (if you begin with one type, you must end with the same type).

Run the code below for an example of the data types! *Note that R is case sensitive! Try running my_logical with lowercase FALSE to see what happens!*

```{r}
# Change my_numeric to be 42
my_numeric <- 42

# Change my_character to be "universe"
my_character <- "universe"

# Change my_logical to be FALSE
my_logical <- FALSE

# Display all three examples
my_numeric
my_character
my_logical
```

If you are ever confused about what data type you are working with, you can actually ask R to help you by calling the function "class". The format is **`class`**`(insert_variable)`. Before you run the code, make a guess what you think the data type will be. Feel free to play around and change stuff in here if you want!

```{r}
# Declare variables of different types
first_variable <- TRUE
second_variable <- 20
third_variable <- "20"

# Check class of my_numeric
class(first_variable)

# Check class of my_character
class(second_variable)

# Check class of my_logical
class(third_variable)
```

**Congratulations! This was the last exercise for this chapter. Head over to the next chapter to get immersed in the world of vectors!**
